% election.tex
% !TeX spellcheck = en_AU
\documentclass[dareport.tex]{subfiles}

\begin{document}
% Content here
\section{Leader Election} \label{sec:election}
\subsection{What is Leader Election?}
A leader or a coordinator in a distributed system is a designated node which has authority over the system\footnote{Through out this text, leader and coordinator are used interchangeably.}. Typical responsibilities of a leader in such a system are mutual exclusion, synchronisation, scheduling, membership coordination, load balancing, etc. In a distributed system, a non-concrete leader is elected in order to coordinate such activities to manage the flawless execution of the processes/servers. An algorithm which is used to elect a leader is known as a leader election algorithm.

The idea of an election is common in real world. A set of entities that has the ability to talk to each other, select a single entity out of them as the leader after some discussion among themselves. When the election process has finished, only one entity remains as the leader, and every other entity agrees and knows the identity of the leader. In a distributed system, conceptually the election process is similar in nature. A distributed system consists of nodes (servers/processes) which are connected via a network which is used by the nodes to communicate. Similar to the real world scenario, once a leader is elected through a leader election algorithm, all of the servers have to agree on that choice. If the leader fails or retires at some point, a new election is conducted to elect a new leader. The leader election policy depends on the algorithm and the requirements of the system.

There are different types of leader election algorithms. Chang and Roberts have introduced a ring based leader election algorithm\cite{ChangRoberts-ringelection} where all the nodes in the system form a logical ring and sends a token in one direction of the ring. The token contains the highest id of the participating nodes and if the current recipient's id is higher than the id in the token, it overwrites the value in the token. When the tokens is moved around the ring and reaches the same node as the id in the token, it means that token has travelled around the ring and contains the highest id in the whole ring and thus that node becomes the leader. One of the main disadvantages of this algorithm is that it does not tolerate failures. When a node has failed the logical rind structure is broken and hence the election process does not work. Therefor it has limited practical value. Another famous leader election algorithm is the Bully Election Algorithm proposed by Garcia-Molina in 1982 \cite{election}. In bully election algorithm every node is aware of other nodes. They elect a node by passing the messages around with each other until the node with the highest priority bullies the others to accept it as the leader. Bully Election Algorithm tolerates failures. It requires $ O(n^2) $ messages to be exchanged to elect a leader compared to $ O(n) $ messages required for the ring election algorithm. However since in real world distributed systems, node failures are common, we have decided to use the Bully Algorithm and an efficient form of the bully algorithm known as the Fast Bully Algorithm\cite{fastbully} for leader election in Strike. 

For Strike project we have implemented the above mentioned two election algorithms, Bully Election and Fast Bully election, to elect a coordinator for our distributed system. Only one of these algorithms could be executed at a time, and is controlled by a configuration value, $election.fast.bully=\{true, false\}$. The responsibilities of the elected leader are explained further in \Cref{sec:consensus}.

\subsection{Features of Leader Election}
Each process in the distributed system has the ability to execute the leader election procedure. However it is only executed if the current leader fails or when a node detects that there's no leader in the system. Therefore in a given multiple processes could be calling on elections. However the elected leader has to be the same for all elections. A leader election algorithm must meet the following two properties \cite{coulouris}.

Each process $p_i$ in the system has a variable $elected_i$, which contains an identifier of the elected process. Initially or when a process is entering the election process, it sets its own $elected$ variable to $\perp$ or \emph{undefined} state. So for each process $i$,

\begin{equation*}
	elected_i =
		\begin{cases}
			\text{P} &: \text{if a process was elected}\\
			\text{$\perp$} &: \text{election was unsuccessful}
		\end{cases}
\end{equation*}

During and after any execution of the algorithm, the following properties must be met.
\begin{enumerate}
\item Safety : Each correct process $p_i$ has $elected_i = P$ or $elected_i = \perp$. At the end of execution only one process $P$ is chosen based on the leader election policy (eg: priority of the processes) of the algorithm.
\item Liveness : After the election all the process either chose a leader or crash.
\end{enumerate}

The leader election policy depends on the requirements of the distributed system. For an example a system may chose to select leader based on priority where a higher priority is given to a process with lowest CPU usage. In Strike, the leader election policy is based on priority of the servers where each server is given a unique integer value during server configuration.
\subsection{Bully Algorithm} \label{ssec:bully}
In this section we describe how we have implemented the Bully Algorithm, complexity of the Bully Algorithm and what are the pros and cons of the Bully Algorithm. The Bully Algorithm\cite{election} is proposed in 1982 as a form of a leader election algorithm to be used in a scenario where one or more processes (possibly including the current leader) has failed. Therefore unlike many election algorithms, Bully Algorithm is able to function even with reasonable process failures.

However the original authors of the Bully Algorithm requires a set of assumptions regarding the nodes and the network\cite{election}.
\begin{enumerate}\label{ba-assumptions}
	\item The network is fully connected.\label{assumption:1}
	\item Each node in the network has a priority that can be clearly compared which each other.\label{assumption:2}
	\item Each node in the network is aware of the priority of every node in the network (a priori knowledge).\label{assumption:3}
	\item The network is synchronised and the delivery of messages is guaranteed.\label{assumption:4}
	\item A process does not know, in advance, whether another process has failed or not.\label{assumption:5}
	\item The communication subsystem, does not spontaneously generate messages.\label{assumption:6}
	\item Each process has access to a persistent storage, where the server states are stored.\label{assumption:7}
	\item There are no transmission errors.\label{assumption:8}
	\item A process never pauses and always responds to incoming messages with no delay.\label{assumption:9}
	\item A process may fail and when it fails it immediately halts all operations. A failure may occur even during an election.\label{assumption:10}
\end{enumerate}
There are three message types exchanged in Bully Algorithm. In Strike these message types available through the \emph{Protocol}.
\begin{itemize}\label{ba-message-types}
	\item \emph{election} : the request message sent by a process to start the election procedure\\
	e.g. : \verb|{"type" : "election", "serverid" : "victor"}|
	\item \emph{answer} : the response message to the \emph{election} message\\
	e.g. : \verb|{"type" : "answer", "serverid" : "victor"}|
	\item \emph{coordinator} : the message which claims that the sender is the new coordinator\\
	e.g. : \verb|{"type" : "coordinator", "serverid" : "victor"}|
\end{itemize}
The \Cref{bully-algorithm} illustrates the pseudocode of the Bully Algorithm (BA). The steps of this algorithm and how it is implemented in Strike is explained below.
\begin{enumerate}[leftmargin=*]
	\item First step is to configure Strike to use the BA for leader election and to set up the required timeout values accordingly. To enable the Standard Bully Algorithm, the configuration value $election.fast.bully=false$ has to be set.
	\item Next the timeout configurations required for the BA should be set. In Strike T1 value is set through a combination of configuration values $ alive.interval $ which means gossip interval in T seconds; and $ alive.error.factor $ which provides the factor by which the $ alive.interval $ is multiplied. So $ T1= alive.interval \times alive.error.factor $. Timeout values T2 and T3 in seconds, should be set respectively via $ election.answer.timeout $ and $ election.coordinator.timeout $. T1, T2, and T3 values should be the same for all the processes.
	\item When a server detects that the current coordinator has not responded within \emph{T1} seconds, it sends an \emph{election} message to every server with a priority higher than itself. And then it begins waiting for $ T2 $ seconds for an \emph{answer} message from another server.
	\item If the server did not receive any \emph{answer} message from another server, then it means there are no higher priority servers in the current system (Assumption \ref{assumption:9}). Therefor it appoints itself as the current coordinator and sends a \emph{coordinator} message to every server with a priority lower than itself. And then stops the election procedure.
	\item If an \emph{answer} message was received from one or more servers, the server begins waiting for \emph{T3} seconds for a \emph{coordinator} message from one of those servers.
	\item If a \emph{coordinator} message has received within \emph{T3} seconds, then it means a server with a higher priority has already elected itself as the current leader. Therefor the server admits the new coordinator and stops the election procedure.
	\item If no \emph{coordinator} message is received within \emph{T3} seconds, then it means one or more higher priority servers that previously sent an \emph{answer} message has failed during the election. Therefor the server restarts the election procedure.
	\item A process may receive an \emph{election} message from a server with a priority lower than itself. This means the sender has detected that the current leader in the system has failed and a new leader has to be elected. Then the recipient server will send an \emph{answer} message to the sender. And then it begins its own election procedure (Steps 3 to 7).
	\item A process may receive a \emph{coordinator} message from a server with a priority higher than itself. This means a new coordinator has been elected in the system. Therefore the recipient admits the new coordinator and stops any election procedure it is running.
	\item When a server is starting, for every server the current coordinator is undefined. After the startup, when the configurations are loaded and the connections have been set up each process will run the leader election procedure (Steps 3 to 7) proposing itself as the leader.
\end{enumerate}

\begin{algorithm}[H]
	\caption{Bully Election Algorithm}
	\label{bully-algorithm}
	\begin{algorithmic}[1]
		\BState \emph{BEGIN}
		\Procedure{StartElection}{$P_{i}$}
		\For {every process $P_{j} : priority(P_{j}) > priority(P_{i}$)}
			\State $P_{i}$ sends \emph{election} message to $P_{j}$
		\EndFor
		\State $P_{i}$ waits for \emph{answer} message for the interval T2.
		\If {no \emph{answer} message within timeout T2}
		
		\Comment No process with priority higher than $P_{i}$ exists.
		
		\Comment $P_{i}$ is the coordinator.
			\For {every process $P_{j} : priority(P_{j}) < priority(P_{i}$)}
				\State $P_{i}$ sends \emph{coordinator} message to $P_{j}$
			\EndFor
			\State $P_{i}$ stops election procedure.
		\Else
		
		\Comment A higher priority process has replied.
			\State $P_{i}$ waits for \emph{coordinator} message for the interval T3.
		\If {$P_{i}$ receives \emph{coordinator} message from $P_{j}$ within timeout T3}
		
		\Comment A higher priority process has elected itself as the leader.
			\State $P_{i}$ admits $P_{j}$ as the new coordinator.
			\State $P_{i}$ stops election procedure.
		\Else
			\State \textproc{StartElection($P_{i}$)}
			\Comment $P_{i}$ restarts the election
		\EndIf
		\EndIf
		\EndProcedure
		\State A process $P_{i}$ does not receive a response within T1 from the coordinator.
		
		\Comment Current coordinator has failed.
		\State\indent \textproc{StartElection($P_{i}$)}
		\State A process $P_{j} \left[priority(j) > priority(i)\right]$ receives an \emph{election} message from $P_{i}$
		\State\indent $P_{j}$ sends an \emph{answer} message to $P_{i}$
		\State\indent \textproc{StartElection($P_{j}$)}
		\State A process $P_{j} \left[priority(j) < priority(i)\right]$ receives a \emph{coordinator} message from $P_{i}$
		
		\Comment A higher priority process has been elected.
		\State\indent $P_{j}$ admits $P_{i}$ as the new coordinator.
		\State\indent $P_{j}$ stops election procedure.
		\BState \emph{END}
	\end{algorithmic}
\end{algorithm}

\subsubsection{Complexity of the Bully Algorithm}\label{sssec:ba-complexity}
In the best case scenario, only the server with the second highest priority detects that the current coordinator is down and sends an \emph{election} message to the servers with higher priority than itself. But since the only server with a higher priority is the current coordinator who has failed, then after timeout $ T2 $ it can elect itself as the new coordinator and inform all the subordinate servers about the new coordinator. Therefor in the best case only $ N-2 $ messages need to be sent.

However in the worst case, consider the server with the lowest priority detects that the current coordinator has failed. Then it will send an \emph{election} message to every server with a higher priority. And upon receipt of the \emph{election} message, after sending an \emph{answer} message, each of these servers will start their own elections (Step 8) by sending \emph{election} messages to every server with a higher priority. So in this scenario $ N-1 $ servers will begin elections. This cascading effect leads to $ O(n^2) $ messages to be exchanged in the worst case.

\subsection{Fast Bully Algorithm} \label{ssec:fast-bully}
The Fast Bully Algorithm (FBA)\cite{fastbully} was designed to overcome the $ O(n^2) $ message complexity of the Bully Algorithm, and as a result to increase the efficiency of the election procedure. However the roots of the FBA come from BA. Similar to BA, FBA is also fault tolerent. So it can cope with the situations where the processes fail before, during or after the election procedure. The FBA also assumes the same assumptions as BA (\Cref{ba-assumptions}).

In addition to the 3 types of messages used in BA in \Cref{ba-message-types}, FBA uses 3 more messages. So 6 types of messages in total are used in FBA.
\begin{itemize}\label{fba-message-types}
	\item \emph{nomination} : the message sent to the process with the highest number to notify that it is the candidate for the coordinator\\
	e.g. : \verb|{"type" : "nomination", "serverid" : "victor"}|
	\item \emph{iamup} : when a process has recovered from a failure, this message is sent\\
	e.g. : \verb|{"type" : "iamup", "serverid" : "victor"}|
	\item \emph{view} : the response message to the \emph{iamup} message. This message contains details of all the servers currently in the system and a separate entry for the current coordinator of the system.\\
	e.g. :
	\begin{small}
		\begin{verbatim}
		{
		"type":"serverlist", "coordinatorid" : "1",
		"servers": 
		[ 	
		{"serverid":"1", "address":"192.168.0.1", "port":"4444"}, 
		{"serverid":"2", "address":"192.168.0.2", "port":"4444"} 
		]
		}
		\end{verbatim}
	\end{small}
\end{itemize}
The \Cref{fast-bully-algorithm-start-failure-recovery,fast-bully-algorithm-start-election,fast-bully-algorithm} illustrates the pseudocode for FBA. The step-by-step guide on how it is implemented in Strike is explained below.

\begin{enumerate}[leftmargin=*]
	\item First step is to configure Strike to use the FBA for leader election and to set up the required timeout values accordingly. To enable the Standard Bully Algorithm, the configuration value $election.fast.bully=true$ has to be set.
	\item Next the timeout configurations required for the FBA should be set. In Strike T1 value is set through a combination of configuration values $ alive.interval $ which means gossip interval in T seconds; and $ alive.error.factor $ which provides the factor by which the $ alive.interval $ is multiplied. So $ T1= alive.interval \times alive.error.factor $. Timeout values T2 and T3 in seconds, should be set respectively via $ election.answer.timeout $ and $ election.coordinator.timeout $. $T1, T2 $, and $ T3 $ values should be the same for all the processes.
	\item Unlike $ T1, T2 $, and $ T3 $ the timeout value $ T4 $ is a different value for each server. This timeout value can be set via $ election.nomination.timeout $ configuration. The $ T4 $ value should be proportionate to the priority of the server, i.e. the server with the highest priority should expire the timeout $ T4 $ first. Furthermore the $ T4 $ value has no effect on the accuracy of the Fast Bully Algorithm, however it does affect the performance and the response time of the algorithm.
	\item When a server detects that it has not received a response from the current coordinator for $ T1 $ seconds, it sends an \emph{election} message to every server with a priority higher than itself. And then it begins waiting for $ T2 $ seconds for an \emph{answer} message from another server.
	\item If the server did not receive any \emph{answer} message from a higher priority server, then it means that there are no higher priority servers in the current system (Assumption \ref{assumption:9}). Therefor it appoints itself as the new coordinator and sends a \emph{coordinator} message to every server with a priority lower than itself. And then stops the election procedure.
	\item If an \emph{answer} message was received by the server, then server collects all the answer messages for $ T2 $ and determines the highest priority server out of the answering servers. Then the server sends a \emph{nomination} message to that server nominating it as a candidate to become the new leader. Then the sender waits for a \emph{coordinator} message from that server for $ T3 $ seconds.
	\item If a \emph{coordinator} message was received, then it means that the nominated candidate server has accepted the leader nomination and become the new leader. Therefor the server admits the new leader and stops the election procedure.
	\item If no \emph{coordinator} message was received, then it means the previously nominated candidate server has failed. Then the server repeats Step 6 by selecting the next server with the highest priority. If there's no server left to chose, the server restarts the election procedure (Steps 4 to 8).
	\item If a server receives an \emph{election} message from a server with a priority lower than itself, then the server sends an \emph{answer} message to the server. Then it waits for $ T4 $ seconds for a \emph{nomination} or a \emph{coordinator} message.
	\item If a \emph{coordinator} message or \emph{nomination} message was not received within $ T4 $ seconds, then it could mean that the server that was conducting the election has failed. And a leader was not elected. Therefor the server restarts the election procedure.
	\item If a server receives a \emph{nomination} message from a server with a priority lower than itself, then the recipient server accepts the nomination and sends a \emph{coordinator} message to all the servers with a priority lower than itself. Then the server stops the election procedure.
	\item If a server receives a \emph{coordinator} message from a server with a priority higher than itself, then it means a new leader has been elected and the recipient server admits the new leader and stops the election procedure.
	\item When a server recovers from a failure or when a server is starting up, it sends an \emph{iamup} message to all the servers. And then waits for a \emph{view} message for $ T2 $ seconds.
	\item If a \emph{view} message was not received within $ T2 $ seconds, it means no other server is up in the system. Therefor the sender becomes the coordinator immediately and stops the election procedure.
	\item If a \emph{view} message was received within $ T2 $ seconds, it compares the received view with its own view. If the views are different, it updates its view i.e it adds the new servers to its servers list. Then if the priority of the current coordinator in the system is higher than its own priority it admits the current coordinator and stops the election procedure. Otherwise, it becomes the new coordinator and sends a \emph{coordinator} message to all the servers with a lower priority and stops the election procedure.
	\item If an \emph{iamup} message is received by a server, it sends its servers list and the current coordinator as a \emph{view} message to the sender.
\end{enumerate}

\begin{algorithm}[h]
	\caption{Fast Bully Election Algorithm - Start failure recovery}
	\label{fast-bully-algorithm-start-failure-recovery}
	\begin{algorithmic}[1]
		\BState \emph{BEGIN}
		\Procedure{StartFailureRecovery}{$P_{i}$}
		\State $P_{i}$ sends \emph{iamup} message to every process
		\State $P_{i}$ waits for \emph{view} message for interval T2
		\If{no \emph{view} message within T2}
		
		\Comment $P_{i}$ is the coordinator
		\State Stop the procedure
		\Else
		\State $P_{i}$ compares the received views with its view.
		\If {view of $P_{i}$ $\ne$ receieved view}
			\State $P_{i}$ updates its view.
		\EndIf
		\If {$priorty(P_{i}) > priority(P_{j})$ $\forall P_{j}$ in its view}
		
		\Comment $P_{i}$ is the highest priority process in the current network.
			\For {every process $P_{j} : priority(P_{j}) < priority(P_{i}$)}
				\State $P_{i}$ sends \emph{coordinator} message to $P_{j}$
				\State $P_{i}$ stops the procedure.
			\EndFor
		\Else
			\State $P_{i}$ admits the highest priority process in its view as the coordinator.
			\State $P_{i}$ stops the procedure.
		\EndIf
		\EndIf
		\EndProcedure
		\BState \emph{END}
	\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
	\caption{Fast Bully Election Algorithm -  Start election}
	\label{fast-bully-algorithm-start-election}
	\begin{algorithmic}[1]
		\BState \emph{BEGIN}
		\Procedure{StartElection}{$P_{i}$}
			\For {every process $P_{j} : priority(P_{j}) > priority(P_{i}$)}
				\State $P_{i}$ sends \emph{election} message to $P_{j}$
			\EndFor
			\State $P_{i}$ waits for \emph{answer} message for the interval T2.
			\If {no \emph{answer} message within timeout T2}
				
			\Comment No process with priority higher than $P_{i}$ exists.
				
			\Comment $P_{i}$ is the coordinator.
				\For {every process $P_{j} : priority(P_{j}) < priority(P_{i}$)}
					\State $P_{i}$ sends \emph{coordinator} message to $P_{j}$
				\EndFor
				\State $P_{i}$ stops election procedure.
			\Else
			
			\Comment Processes with higher priority have answered.
				\State add answered process to answered list.
				\State send \emph{nomination} message to $P_{j}$ : $priority(P_{j}) > priority(P_{k})$ $\forall P_{k}$ that answered  $P_{i}$ \label{send-nomination}
				\State $P_{i}$ waits for \emph{coordinator} message for interval T3.
				\If {A coordinator message from $P_{j}$ is received}
					\State $P_{i}$ admits $P_{j}$ as the new coordinator.
					\State $P_{i}$ stops election procedure.
				\Else
					\State Remove $P_{j}$ from answered list.
					\State Repeat \cref{send-nomination} for the next $P_{j}$ in the answered list.
					\If {Answered list is empty}
						\State \textproc{StartElection($P_{j}$)}
					\EndIf
				\EndIf
			\EndIf
		\EndProcedure
		\BState \emph{END}
	\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
	\caption{Fast Bully Election Algorithm}
	\label{fast-bully-algorithm}
	\begin{algorithmic}[1]
		\BState \emph{BEGIN}
		\State A process $P_{i}$ recovers from failure
			\State\indent\textproc{StartFailureRecovery($P_{i}$)}
		\State A process $P_{i}$ may receive an IamUp message from $P_{j}$
			\State\indent $P_{i}$ sends view message to $P_{j}$.
		\State A process $P_{i}$ does not receive a response within T1 from the coordinator
			\State\indent\textproc{StartElection($P_{i}$)}
		\State A process $P_{j} \left[priority(j) > priority(i)\right]$ receives an \emph{election} message from $P_{i}$
			\State\indent $P_{j}$ sends an \emph{answer} message to process $P_{i}$
			\State\indent $P_{j}$ waits for either a \emph{coordinator} or a \emph{nomination} message for interval T4
			\State\indent {\textbf{if} no \emph{coordinator} or \emph{nomination} message within T4 \textbf{then}}
				\State\indent\indent\textproc{StartElection($P_{j}$)}
			
			\State { A process $P_{j} \left[priority(j) > priority(i)\right]$ receives a \emph{nomination} message from $P_{i}$}
			
			\Comment The current election manager has decided $P_{j}$ to be the new coordinator\footnote{Current election manager is the process which started the ongoing election}.
				\State\indent $P_{j}$ admits itself as the new coordinator.
				\State\indent {\textbf{for} every process $P_{k} : priority(P_{k}) < priority(P_{j}$) \textbf{do}}
					\State\indent\indent $P_{j}$ sends \emph{coordinator} message to $P_{k}$
					\State\indent\indent $P_{j}$ stops the procedure.
			\State {A process $P_{j} \left[priority(j) < priority(i)\right]$ receives a \emph{coordinator} message from $P_{i}$}
			
			\Comment A higher priority process has been elected.
				\State\indent $P_{j}$ admits $P_{i}$ as the new coordinator.
				\State\indent $P_{j}$ stops election procedure.
		\BState \emph{END}
	\end{algorithmic}
\end{algorithm}
\subsubsection{Complexity of the Fast Bully Algorithm}
The best case scenario for the Fast Bully Algorithm is the same as the best case scenario for the Bully Algorithm (\Cref{sssec:ba-complexity}). However in the worst case scenario, FBA has a significant performance increase in terms of the number of the messages. In the worst case scenario, the server with the lowest priority (say $ P_w $)detects the current coordinator has failed and initiates the election algorithm. It sends \emph{election} messages to $ N-2 $ servers and receives $ N-2 $ \emph{answer} messages. Then it selects the highest priority server out of them and sends a $ 1 $ \emph{nomination} message. The nominated server accepts the nomination and sends $ N-2 $ \emph{coordinator} messages. Therefor totally only $ O(n) $ messages are required for FBA.
\subsection{How Leader Election is implemented in Strike}
In Strike, we have set up a gossip based reliable failure detector to initiate the leader election in case the leader is detected to be failed. For the leader we do not go through the consensus process since the leader is required (explained more in \Cref{sec:consensus}) to initiate the consensus process. If the leader is suspected to have failed, the detected server will immediately start the election process. This has two advantages. First one is if the suspected leader has really failed, it allows a new leader to be elected immediately. Second advantage is if the suspected leader has not failed, it can participate in the leader election procedure and make sure that it is reelected as the leader. Therefor always the correct leader is elected. In addition to that, when a server is starting it will start either the \textproc{StartElection($P_{i}$)} procedure of the \Cref{bully-algorithm} (if BA is enabled) or the \textproc{StartFailureRecovery($P_{i}$)} procedure of the \Cref{fast-bully-algorithm} (if FBA is enabled) to make sure the newly joining server do not affect the correctness of the elected leader as per the leader election policy. For Strike the leader election policy is based on the priority of the servers. Each server is given an integer value (eg: 1, 2, 3. \dots) as an identifier and the server with the lowest integer value has the highest priority. The leader election algorithm uses this policy to elect the leader when multiple servers participate in the leader election. For an example if 3 servers with identifiers $ 1, 2, 3 $ respectively then server with the identifier $ 1 $ will become the leader after the leader election process.

\subsection{Discussion}
The Bully algorithm is a widely used election algorithm due to its fault tolerance capabilities and simplicity. The Fast Bully Algorithm is a modification to the Bully Algorithm to increase the performance of it in terms of reduced number of messages and reduced response time. The number of messages has been reduced from $ O(n^2) $ to $ O(n) $ messages while the response time has been reduced by 35 to 80\% depending on the scenario. Even though the FBA is complex and higher in implementation cost, it is not very difficult to implement the FBA by utilising a similar flow and architecture to an implementation of BA. In contrast to the 3 message types in BA, FBA uses 6 message types (3 message types in addition to the message types used in BA). Therefor the performance gain that can be achieved by FBA outweighs the implementation overhead. We were able to observe a significant increase in the response time with FBA specially when a higher priority server joins the network, it immediately becomes the new coordinator since it does not have to hold an election to be elected. When the number of Strike servers were increased, the network congestion and the time taken to elect a leader reduced significantly. An additional advantage of these two algorithms is the fact that the number of times the algorithms are executed does not affect the final result. Even if all the servers have executed the [BA or FBA] algorithm at the same time, all the servers will elect the same coordinator at the end.

One of the main drawbacks of BA is the set of assumptions it depends on, where in real world systems these assumptions may not hold. For an example the Assumption \ref{assumption:9} (\Cref{ba-assumptions}) expects the servers to immediately reply when it receives a message. However in real systems, due to the performance issues at the server (server could be overloaded to process the incoming messages) and the transmission delays, the message processing and transmission could get delayed. And since BA relies on timeouts to move between states, these delays could affect the accuracy of the algorithms. One of the solutions for this issue is to detect late messages and to restart the election procedure in order to make sure that a correct coordinator is elected. However with BA, this is an expensive task. Even though the FBA also depends on the same set of assumptions as BA, it provides an easy procedure to execute the solution to this problem. When a server detects a late message or when it detects that it has recovered from a failure, the server can execute the \textproc{StartFailureRecovery($P_{i}$)} procedure of the \Cref{fast-bully-algorithm-start-failure-recovery} to make sure the whole system has a correct coordinator. This can be executed by exchanging $ O(n) $ messages as opposed to $ O(n^2) $ messages required in BA. However, even though the algorithms can be implemented to recover from a failure, it requires a reliable failure detector to make sure the it is executed correctly and appropriately.

In conclusion, it is often required to have a coordinator or a leader in a distributed system to ensure that some tasks are executed flawlessly. In Strike we use the leader to conduct the consensus process. Both the Bully Algorithm and the Fast Bully Algorithm are leader election algorithms which are strong in operating in an environment where nodes could fail, and in real world distributed systems this happens often. Since Fast Bully Algorithm is an incremental improvement on top of Bully Algorithm, it has increased performance while some of the disadvantages of BA are mitigated.
\end{document}